Optimal updating factor in Wang-Landau and metadynamics simulations

### 2016-04-06 doc/invt.tex
predict.c: merging the scan code

For the single-bin updating scheme,
we chose $\alpha_0 = 0.0001$
and ran the simulation for $t = 10^8$ steps.
%
As shown in Fig. \ref{fig:err_singlebin},
the numerical results of the terminal error agreed well with
the analytical prediction by Eq. \eqref{eq:error_invt}.
%
For both global and local sampling processes,
the optimal $\lambda$ occurred around $1.0$,
verifying Eq. \eqref{eq:alpha_invt1}.
%
%Recall that $1/\lambda$ in Eq. \eqref{eq:alpha_invtlambda}
%represents the relative updating magnitude.
%%
%Figure \ref{fig:err_singlebin} also shows
%an asymmetry of the error curve.
%A smaller than optimal value of
%the updating magnitude, $1/\lambda$,
%tends to increase the error more rapidly
%than a larger than optimal value,
%suggesting that the over-updating
%is generally preferable to the under-updating
%when the optimal value is unknown.

The two error functions
differed only by a multiplicative constant,
again agreeing with the theory.
%
With identical eigenvalues,
$\lambda_1 = \cdots = \lambda_n = 1$,
the error in Eq. \eqref{eq:error_invt}
depends on the sampling process only through
the sum $\sum_{ k = 2 }^n \Gamma_k$,
which gives rise to the above multiplicative constant.

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsbin.pdf}
  }
  \caption{
    \label{fig:err_singlebin}
    Error, $E$, versus the
    relative updating magnitude,
    %proportionality constant,
    $1/\lambda$,
    in Eq. \eqref{eq:alpha_invtlambda}
    for the single-bin (WL) scheme.
    %
    The results have been averaged over $1000$ independent runs.
  }
\end{center}
\end{figure}

>>>>

For the single-bin updating scheme,
we chose $\alpha_0 = 0.0001$
and ran the simulation for $t = 10^8$ steps.
%
As shown in Fig. \ref{fig:errsbin},
the numerical results of the terminal error agreed well with
the analytical prediction by Eq. \eqref{eq:error_invt}.
%
For both global and local sampling processes,
the optimal $\lambda$ occurred around $1.0$,
verifying Eq. \eqref{eq:alpha_invt1}.
%
%Recall that $1/\lambda$ in Eq. \eqref{eq:alpha_invtlambda}
%represents the relative updating magnitude.
%%
%Figure \ref{fig:errsbin} also shows
%an asymmetry of the error curve.
%A smaller than optimal value of
%the updating magnitude, $1/\lambda$,
%tends to increase the error more rapidly
%than a larger than optimal value,
%suggesting that the over-updating
%is generally preferable to the under-updating
%when the optimal value is unknown.
%
The two error functions
differed only by a multiplicative constant,
again agreeing with the theory.
%
With identical eigenvalues,
$\lambda_1 = \cdots = \lambda_n = 1$,
the error in Eq. \eqref{eq:error_invt}
depends on the sampling process only through
the sum $\sum_{ k = 2 }^n \Gamma_k$,
which gives rise to the above multiplicative constant.

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsbin.pdf}
  }
  \caption{
    \label{fig:errsbin}
    Error, $E$, versus the
    relative updating magnitude,
    %proportionality constant,
    $1/\lambda$,
    in Eq. \eqref{eq:alpha_invtlambda}
    for the single-bin (WL) scheme.
    %
    The results have been averaged over $1000$ independent runs.
    %
    \note{The figure is produced by \texttt{doc/fig/errsbin.gp}.
      The data are produced by the script \texttt{invt2run.py},
      and saved under \texttt{data/singlebin}
      (see the \texttt{Makefile} there).
    }
  }
\end{center}
\end{figure}
